Comparison of dimensionality reduction schemes for parallel global optimization algorithms

### 2018-05-10 2018/YSC/ysc_paper.tex
YSC: add full version of paper

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%%%%%
\usepackage{eqnarray,amsmath}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amssymb}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}
\setlength{\textfloatsep}{4pt plus 1.0pt minus 1.5pt}
\setlength{\floatsep}{4pt plus 1.0pt minus 1.5pt}

\begin{document}
\setlength{\abovedisplayskip}{3pt}
\setlength{\belowdisplayskip}{3pt}
%
\title{Comparison of dimensionality reduction schemes for parallel global optimization
algorithms 
\thanks{The study was supported by the Russian Science Foundation, project No 16-11-
10150}}
%
\titlerunning{Dimension reduction schemes in parallel GO}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Konstantin Barkalov \and
Vladislav Sovrasov \and
Ilya Lebedev}
%
\authorrunning{K. Barkalov et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{
Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia
\email{konstantin.barkalov@itmm.unn.ru}\\
\email{sovrasov.vlad@gmail.com}\\
\email{ilya.lebedev@itmm.unn.ru}\\
\url{http://hpc-education.unn.ru/}
}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
This work considers a parallel algorithms for solving multi-extremal optimization problems.
Algorithms are developed within the framework of the information-statistical approach and
implemented in a parallel solver Globalizer. The optimization problem is solved by reducing
the multidimensional problem to a set of joint one-dimensional problems that are solved in
parallel. Five types of Peano-type space-filling curves are employed to reduce dimension. The
results of computational experiments carried out on several hundred test problems are discussed.

\keywords{Global optimization \and  Dimension reduction \and Parallel algorithms  \and
Multidimensional multiextremal optimization  \and Global search algorithms  \and Parallel
computations }
\end{abstract}

%------------------------------------------------------------------------------
\section{Introduction}\label{sec:intro}
%\begin{Russian}

@@

%\end{Russian}
%------------------------------------------------------------------------------
\section{Statement of Multidimensional Global Optimization Problem}
In this paper, the core class of optimization problems, which can be solved using
Globalizer, is formulated. This class involves the multidimensional global
optimization problems without constraints, which can be defined in the following way:
\begin{equation}
\label{eq:task}
\begin{array}{cr}\\
  \varphi(y^*)=\min\{\varphi(y):y\in D\}, \\
  D=\{y\in \mathbb{R}^N:a_i\leq y_i\leq{b_i}, 1\leq{i}\leq{N}\}
\end{array}
\end{equation}
with the given boundary vectors  $a$ and  $b$. It is supposed, that the objective function
\(\varphi(y)\) satisfies the Lipschitz condition
\begin{equation}
\label{eq:lip}
|\varphi(y_1)-\varphi(y_2)|\leq L\Vert y_1-y_2\Vert,y_1,y_2\in D,
\end{equation}
where \(L>0\) is the Lipschitz constant, and \(||\cdot||\) denotes the norm in \(\mathbb{R}^N\)
space.
\par
Usually, the objective function \(\varphi(y)\) is defined as a computational procedure,
according to which the value \(\varphi(y)\) can be calculated for any vector \(y\in D\)
(let us further call such a calculation \textit{a trial}). It is supposed that this procedure
is time-consuming.
%As a result, the overall time of solving the optimization
%problem (\ref{eq:task}) is determined, first of all by the number of executed trials.
%It should also be noted that the requirement of the Lipschitz condition (\ref{eq:lip})
%is highly important, since an estimate of the global minimum can be constructed on the
%basis of a finite number of computed values of the optimized function only in this case .

@@

In Sec.~\ref{sec:results_parallel} the results of the numerical experiments are presented, which
have demonstrated the algorithm from Sec.~\ref{sec:parallel_evolvents} based on the rotated
evolvents allowed obtaining the speedup up to 43 times when solving the problem series
employing several nodes of the computer cluster. It is worth noting that the objective functions
in the considered problems are not computation-costly (the averaged computation time was
$10^{-3}$ sec). In the case of more complex problems,the speedup in time could approach the
speedup with respect to the number of iterations.
%\end{Russian}

% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\begin{thebibliography}{107}

@@

\bibitem{Evtushenko}
Evtushenko, Y., Posypkin, M.: \emph{A deterministic approach to global box-constrained
optimization.} Optim. Lett. \textbf{7(4)}, 819--829 (2013)

\end{thebibliography}
\end{document}

>>>>

% easychair.tex,v 3.1 2011/12/30
%
% Select appropriate paper format in your document class as
% instructed by your conference organizers. Only withtimes
% and notimes can be used in proceedings created by EasyChair
%
% The available formats are 'letterpaper' and 'a4paper' with
% the former being the default if omitted as in the example
% below.
%
\documentclass[procedia]{easychair}
%\documentclass[debug]{easychair}
%\documentclass[verbose]{easychair}
%\documentclass[notimes]{easychair}
%\documentclass[withtimes]{easychair}
%\documentclass[a4paper]{easychair}
%\documentclass[letterpaper]{easychair}

% This provides the \BibTeX macro
\usepackage{doc}
\usepackage{makeidx}

\usepackage{eqnarray,amsmath}
\usepackage{subfig}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amssymb}

\setlength{\textfloatsep}{4pt plus 1.0pt minus 1.5pt}
\setlength{\floatsep}{4pt plus 1.0pt minus 1.5pt}

% In order to save space or manage large tables or figures in a
% landcape-like text, you can use the rotating and pdflscape
% packages. Uncomment the desired from the below.
%
% \usepackage{rotating}
% \usepackage{pdflscape}

% If you plan on including some algorithm specification, we recommend
% the below package. Read more details on the custom options of the
% package documentation.
%
% \usepackage{algorithm2e}

\def\procediaConference{99th Conference on Topics of
  Superb Significance (COOL 2014)}

%\makeindex

%% Front Matter
%%
% Regular title as in the article class.
%
\title{Comparison of dimensionality reduction schemes for parallel global optimization
algorithms}

% \titlerunning{} has to be set to either the main title or its shorter
% version for the running heads. When processed by
% EasyChair, this command is mandatory: a document without \titlerunning
% will be rejected by EasyChair

\titlerunning{Dimension reduction schemes in parallel GO}

% Authors are joined by \and. Their affiliations are given by \inst, which indexes into the list
% defined using \institute
%
\author {
Vladislav Sovrasov \and
Ilya Lebedev
}

% Institutes for affiliations are also joined by \and,
\institute  {
  Lobachevsky State University of Nizhni Novgorod, Nizhni Novgorod, Russia
  \email{sovrasov.vlad@gmail.com}\\
  \email{ilya.lebedev@itmm.unn.ru}\\
}
%  \authorrunning{} has to be set for the shorter version of the authors' names;
% otherwise a warning will be rendered in the running heads. When processed by
% EasyChair, this command is mandatory: a document without \authorrunning
% will be rejected by EasyChair

\authorrunning{Sovrasov V., Lebedev I.}

\begin{document}

\maketitle

\keywords{Global optimization, Dimension reduction, Parallel algorithms,
Multidimensional multiextremal optimization, Global search algorithms}

\begin{abstract}
This work considers a parallel algorithms for solving multi-extremal optimization problems.
Algorithms are developed within the framework of the information-statistical approach and
implemented in a parallel solver Globalizer. The optimization problem is solved by reducing
the multidimensional problem to a set of joint one-dimensional problems that are solved in
parallel. Five types of Peano-type space-filling curves are employed to reduce dimension. The
results of computational experiments carried out on several hundred test problems are discussed.
\end{abstract}

%\setcounter{tocdepth}{2}
%{\small
%\tableofcontents}

%\section{To mention}
%
%Processing in EasyChair - number of pages.
%
%Examples of how EasyChair processes papers. Caveats (replacement of EC
%class, errors).

%------------------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}

@@

%------------------------------------------------------------------------------
\section{Statement of Multidimensional Global Optimization Problem}
In this paper, the core class of optimization problems, which can be solved using
Globalizer, is formulated. This class involves the multidimensional global
optimization problems without constraints, which can be defined in the following way:
\begin{equation}
\label{eq:task}
\begin{array}{cr}\\
  \varphi(y^*)=\min\{\varphi(y):y\in D\}, \\
  D=\{y\in \mathbb{R}^N:a_i\leq y_i\leq{b_i}, 1\leq{i}\leq{N}\}
\end{array}
\end{equation}
with the given boundary vectors  $a$ and  $b$. It is supposed, that the objective function
\(\varphi(y)\) satisfies the Lipschitz condition
\begin{equation}
\label{eq:lip}
|\varphi(y_1)-\varphi(y_2)|\leq L\Vert y_1-y_2\Vert,y_1,y_2\in D,
\end{equation}
where \(L>0\) is the Lipschitz constant, and \(||\cdot||\) denotes the norm in \(\mathbb{R}^N\)
space.
\par
Usually, the objective function \(\varphi(y)\) is defined as a computational procedure,
according to which the value \(\varphi(y)\) can be calculated for any vector \(y\in D\)
(let us further call such a calculation \textit{a trial}). It is supposed that this procedure
is time-consuming.
%As a result, the overall time of solving the optimization
%problem (\ref{eq:task}) is determined, first of all by the number of executed trials.
%It should also be noted that the requirement of the Lipschitz condition (\ref{eq:lip})
%is highly important, since an estimate of the global minimum can be constructed on the
%basis of a finite number of computed values of the optimized function only in this case .

@@

In Sec.~\ref{sec:results_parallel} the results of the numerical experiments are presented, which
have demonstrated the algorithm from Sec.~\ref{sec:parallel_evolvents} based on the rotated
evolvents allowed obtaining the speedup up to 43 times when solving the problem series
employing several nodes of the computer cluster. It is worth noting that the objective functions
in the considered problems are not computation-costly (the averaged computation time was
$10^{-3}$ sec). In the case of more complex problems,the speedup in time could approach the
speedup with respect to the number of iterations.
%------------------------------------------------------------------------------
% Index
%\printindex
\begin{thebibliography}{107}

@@

  \bibitem{Evtushenko}
  Evtushenko, Y., Posypkin, M.: \emph{A deterministic approach to global box-constrained
  optimization.} Optim. Lett. \textbf{7(4)}, 819--829 (2013)
\end{thebibliography}

%------------------------------------------------------------------------------
\end{document}

% EOF
